---
title: "A5 Q8.6"
author: "Joshua Levine"
output: html_document
date: "2026-02-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question
In R, what’s the difference between `[`, `[[`, and `$` when applied to a list?

## Answer
To answer this question, we will create a list and play with the three operators.
```{r}
my_list <- list(ints = 2:5, 1, logicals = c(TRUE, FALSE))
my_list[1]
typeof(my_list[1])
typeof(my_list[2])
names(my_list[1])
my_list[c(1, 2)]
my_list[[1]]
typeof(my_list[[1]])
names(my_list[[1]])
my_list[[c(1, 2)]]
my_list$ints
typeof(my_list$ints)
names(my_list$ints)
```

We draw the following conclusions:

* `[` will extract the element(s) of a list corresponding to the inputted indices. The extracted elements are always returned as lists, regardless of their respective data types and whether they have names.
* `[[` can only extract a single "immediate" element of the list at a time, but will return it in its "default" data type (outside of a list) while dropping its name.
* `$` is a convenient version of `[[` for named elements of a list, as the extracted element is referred to by its name rather than its index.